---
title: "Homework 2 Starter Code"
author: "Richard Chen"
date: "April 8, 2016"
output: pdf_document
---

# Note
This pdf is produced by R Markdown. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. One way to produce nice-looking reports for your homework assignments using R Markdown, you need to install the package \texttt{knitr} <http://yihui.name/knitr/>. This document is prepared by R Markdown which is built on the \texttt{knitr} package aforementioned.

For any R function which you feel confused about, please utilize the help function in R for detailed instructions. For example, function \texttt{cor()} help use to compute correlation (matrices), if you would like to know the details of its usage, input the R command \texttt{help(cor)}.

# Additional note:

1. In the case  where you have installed some package, for example \texttt{knitr}, and relaunch R and need to use some functionalities of the R package , you need to run the command \texttt{library(knitr)} before you use the functionalities of the corresponding package. If you never installed the package before, you need to run \texttt{install.packages("knitr")}, or click the "Packages" button on the top of the right lower window in the R Studio interface.

2. Do not forget to set your working directory at the begining.


# Example R Code for Question 1

## a
Download and read the data
```{r}
# download data to the current working folder
download.file("https://github.com/mlakolar/BUS41000/raw/master/data/CountryMonthlyReturns2.csv",
              destfile="CountryMonthlyReturns2.csv")
# assign the data to a variable
countryReturn_df = read.csv("CountryMonthlyReturns2.csv")
```

Construct the returns of the equi-weighted portfolio:
```{r}
port_df = countryReturn_df[,c("Canada", "Japan")]
port_df$port1 = 0.5*port_df$Canada + 0.5*port_df$Japan
head(port_df)
```
then compute the mean and standard deviation of the return series
```{r}
mean(port_df$port1); sd(port_df$port1)
```

## c
When the weight for the Canada is 25% and 75% respectively,
```{r}
port_df$port2 = 0.25*port_df$Canada + 0.75*port_df$Japan
port_df$port3 = 0.75*port_df$Canada + 0.25*port_df$Japan

mcr = sapply(port_df, mean)            # compute mean for each column
sdcr = sapply(port_df, sd)             # compute standard deviation for each column  
```

Now make the standard deviation-mean scatter plot:
```{r}
plot(sdcr, mcr, type="p", main="Scatter for portfolios", xlab = "std. dev.", ylab = "mean", cex=0.7)
# pointLabel(sdcr, mcr, labels=names(mcr), cex= 1)
```

## d. efficient frontier
```{r}
weights = seq(0,1,0.01)
m = weights*mean(port_df$Canada)+(1-weights)*mean(port_df$Japan)
s = sqrt(weights^2*var(port_df$Canada)+(1-weights)^2*var(port_df$Japan)
               +2*weights*(1-weights)*cov(port_df$Canada,port_df$Japan))
plot(s, m, type="b", col="blue", cex = 0.2,
     xlab="std. dev", ylab="mean", main="Porfolio Frontier")
lines(s,m,type="l", col="blue")
```


# Code for Question 3
Download and read the data (reminder: we saw this dataset in lecture01: see 01_core.R )
```{r}
# download data to the current working folder
download.file("https://github.com/mlakolar/BUS41000/raw/master/data/bank.csv",
              destfile="bank.csv")
# assign the data to a variable
bank = read.csv("bank.csv")
int = bank$InterarrivalTime
```


# Sample Code for Question 4
Download and read the data
```{r}
# download data to the current working folder
download.file("https://github.com/mlakolar/BUS41000/raw/master/data/studenthw.csv",
              destfile="studenthw.csv")
# assign the data to a variable
studenthw = read.csv("studenthw.csv")
```

## b
Scatter plot of heights and weights:
```{r}
head(studenthw)
plot(studenthw$height, studenthw$weight, pch=20,
     xlab = "students' heights (inches)", ylab = "students' weights (pounds)", main = "")
```

## e
Scatter plot of heights in feet and heights in inches
```{r}
height.inch = studenthw$height
height.foot = studenthw$height/12
plot(height.foot, height.inch, pch=20,
     xlab = "heights in feet", ylab = "heights in inches", main = "")
```


# Question 5. First Stab at Linear Regression
Download and read the data
```{r}
# download data to the current working folder
download.file("https://github.com/mlakolar/BUS41000/raw/master/data/housesp1.csv",
              destfile="housesp1.csv")
# assign the data to a variable
homep_df = read.csv("housesp1.csv")
```

## b. linear coefficients by R
The R output about the linear regression estimates is:
```{r}
# creat scatter plot
plot(homep_df$size, homep_df$price, col = "blue", xlab = "Size (sq. feet)", ylab = "Price")

# add a regression linea
reg = lm(price~size, data = homep_df)
abline(reg, lw = 2, col = "red")

# summary of linear regression
summary(reg)
```
```{r}
hist(homep_df$price, breaks=20, 
     main="Histogram of house prices", xlab="Price", ylab="Frequency")
```

## d. prediction baesd on linear regression
```{r}
# download data to the current working folder
download.file("https://github.com/mlakolar/BUS41000/raw/master/data/housesp2.csv",
              destfile="housesp2.csv")
# assign the data to a variable
homep2_df = read.csv("housesp2.csv")

# the new data only contains column sizenew
head(homep_df)

# rename the column, it needs to be the same name as in homep_df
colnames(homep2_df) = c("size")
```
Predict prices for new houses
```{r}
yhat = predict(reg, homep2_df)
head(yhat)
```

# Question 6. Wine Offers

Download data to the current working folder
```{r}
download.file("https://github.com/mlakolar/BUS41000/raw/master/data/Wine_OfferInformation.csv", 
              destfile="Wine_OfferInformation.csv")
download.file("https://github.com/mlakolar/BUS41000/raw/master/data/Wine_Transactions.csv", 
              destfile="Wine_Transactions.csv")

# load data into R
offers_df       = read.csv("Wine_OfferInformation.csv", row.names = 1)
transactions_df = read.csv("Wine_Transactions.csv")
```

## a. cluster analysis
```{r}
# initializa an empty table where rows correspond to customers and columns correspond to offers
customer_by_offer = matrix(0, length(unique(transactions_df[,1])), nrow(offers_df))      
rownames(customer_by_offer) = levels(transactions_df[,1])
colnames(customer_by_offer) = 1:nrow(offers_df)

# put transactions into the table
for (i in 1:nrow(transactions_df)) {
  customer_by_offer[ transactions_df[i,1], transactions_df[i,2] ] = 1  
}

# customer_by_offer now contains information on which offers did each customer get
# for example, here are offers by "Fisher" or "Smith"
customer_by_offer[ "Fisher", ] 
customer_by_offer[ "Smith", ] 

# find 4 clusters
set.seed(1)       # for reproductivity
grpCustomers = kmeans(customer_by_offer, centers = 4, nstart = 1000)

# take a look at cluster means and try to summarize different groups
grpCustomers
```

## b. Wines favored by customers in each cluster
Let check what offers did members of each cluster buy -- rank the offers from most popular of offer to least popular offer within each cluster.
```{r}
# find members in each cluster
cluster.1 = rownames(customer_by_offer)[grpCustomers$cluster==1]
cluster.2 = rownames(customer_by_offer)[grpCustomers$cluster==2]
cluster.3 = rownames(customer_by_offer)[grpCustomers$cluster==3]
cluster.4 = rownames(customer_by_offer)[grpCustomers$cluster==4]

# extract the information of the offers in each cluster
cluster_by_offer = matrix(0,4,nrow(offers_df))
rownames(cluster_by_offer) = 1:4
colnames(cluster_by_offer) = 1:nrow(offers_df)

for (i in 1:nrow(offers_df)){
  cluster_by_offer[1,i] = sum(customer_by_offer[cluster.1,i])
  cluster_by_offer[2,i] = sum(customer_by_offer[cluster.2,i])
  cluster_by_offer[3,i] = sum(customer_by_offer[cluster.3,i])
  cluster_by_offer[4,i] = sum(customer_by_offer[cluster.4,i])
}

# Rank the offers by their numbers in each cluster
sort(cluster_by_offer[1,], decreasing = TRUE)
sort(cluster_by_offer[2,], decreasing = TRUE)
sort(cluster_by_offer[3,], decreasing = TRUE)
sort(cluster_by_offer[4,], decreasing = TRUE)
```
